Post-Disturbance Tree Community Trajectories in a Neotropical Forest

### 2018-06-26 WholePlotTrajectories.tex
comments bubs

\Abstract{
Understanding the ecological rules underlying the maintenance of tropical forest
functioning and dynamics is urgent to anticipate their fate in the
global changing context. The huge diversity if tropical forests is often assumed to be shaped
by constant regime of disturbance yielding a diversity peak at
intermediate intensity, but for tropical forests this intermediate
disturbance hypothesis (IDH) remains debated ????as well as the completeness
their resilience in all taxonomic and functional facets after
disturbance????. To disentangle the ecological processes driving communities
response to disturbance we analysed the taxonomic and functional
diversity trajectories following a logging and thinning disturbance gradient.
Specifically we examined over 30 years the trajectories of
taxonomic richness and evennes and functional composition, diversity and
redundancy based on 7 leaf, stem and life history traits. Trajectories highlighted communities taxonomic
resilience, maintaining their initial differences in composition, and
functional resilience through fast functional trajectory common to all
communities in the functional space. The IDH consistently predicted
communities functional trajectories according to the disturbance
intensity, but poorly represented their taxonomic trajectories, blurried
by the alteration of the functional redundancy and the recruitment of
infrequent, shade tolerant species slowed by competitive exclusion
processes. Although communities functioning had recovered after 30 years
their taxonomic diversity and composition in infrequent species and
their functional redundancy remained altered. these results acknowledged
the need of decades-long recovery cycles to ensure communities complete
resilience, and questioned communities maintenance after repeated
disturbance.
}

@@

In tropical forest, ecological communities are constantly re-shaped by
natural disturbance events that change both the abiotic environment,
through the fluxes of light, heat and water \citep{Goulamoussene2017},
and the biotic interactions like competition among species
\citep{Herault2018}. One of the cornerstone of tropical forest ecology is to
understand the processes and drivers of ecosystems response to
disturbance \citep{White2001, Chazdon2003a}. For now, this has been
largely studied through forest structural parameters, rapid and convenient to
measure, as aboveground biomass, tree height or stem density
\citep{Piponiot2016, Rutishauser2016}. These structural parameters were
thereafter sucessfully modeled, giving important insights into the
recovery of ecosystems processes and services
\citep{Herault2018}. However the response of forests
diversity is still unclear, albeit it determines the
productivity, stability and functioning of ecosystems
\citep[\citet{Liang2016}]{Tilman2014} and would be most probably
impacted by the changes induced by disturbance \citep{Baraloto2012a}.

@@

\begin{table}

\caption{\label{tab:Tab1}Intervention table, summary of the disturbance intensity for the 4 plot treatments in Paracou.}
\centering
\begin{tabu} to \linewidth {>{\raggedright}X>{\raggedright}X>{\raggedright}X>{\raggedright}X>{\raggedright}X}
\toprule
Treatment & Timber & Thinning & Fuelwood & \%AGB lost\\
\midrule
Control &  &  &  & 0\\
T1 & DBH $\geq$ 50 cm, commercial species, $\approx$ 10 trees/ha &  &  & $[12\%-33\%]$\\
T2 & DBH $\geq$ 50 cm, commercial species, $\approx$ 10 trees/ha & DBH $\geq$ 40 cm, non-valuable species, $\approx$ 30 trees/ha &  & $[33\%-56\%]$\\
T3 & DBH $\geq$ 50 cm, commercial species, $\approx$ 10 trees/ha & DBH $\geq$ 50 cm, non-valuable species, $\approx$ 15 trees/ha & 40 cm $\leq$ DBH $\leq$ 50 cm, non-valuable species, $\approx$ 15 trees/ha & $[35\%-66\%]$\\
\bottomrule
\end{tabu}
\end{table}

@@

The study site corresponds to a tropical rainforest typical if the Guiana Shield with a dominance of
Fabaceae, Chrysobalanaceae, Lecythidaceae and Sapotaceae botanical
families. In the twelve experimental plots of the experiment, all trees
above 10 cm DBH have been mapped and measured annually since 1984. Trees are
first identified during inventories with a vernacular name assigned by
the forest worker team, and afterward with a scientific name assigned by
botanists during regular botanical campaigns. In 1984, specific
vernacular names are given to 62 commercial or common species whereas
more infrequent ones were identified under general identifiers only
distinguishing trees and palm trees. From 2003, botanical campaigns have been
conducted every 5 to 6 years to identify all trees at the species level
but identification levels still varied among plots and campaigns.

@@

Six functional traits
representing leaf economics (leaves thickness, toughness, total
chlorophyll content and specific leaf area, the leaf area per unit dry
mass) and wood economics (wood specific gravity and bark thickness), and
life history traits (maximum specific height and seed mass) came from the BRIDGE project \footnote{http://www.ecofog.gf/Bridge/}. Trait values were assessed from a selection of individuals located
in nine permanent plots in French Guiana, including two in Paracou, and
comprised 294 botanical species pertaining to 157 botanical genera.
Missing trait values were filled using multivariate imputation by
chained equation \citep{Mice2011}.
Imputations were restricted within genus, or family when samples were too
scarce, in order to account for the phylogenetic signal of the
functional traits. Whenever a species inventoried was not in the
dataset, it was attributed a set of traits values randomly sampled among
species of the same next higher taxonomic level (same genus or family).
As seed mass information corresponds to a classification into mass
classes, no data filling process was applied and analysis were
restricted to the 414 botanical species recorded.

@@

To counter taxonomic uncertainties due to the variability of botanical
identification levels (in space) and protocols (in time) (see {[}\#protocols{]}), the taxonomic
composition and diversity analysis were conducted at the genus level,
\emph{i.e.} referring to the genus of observed or trialed botanical
names. Trajectories of communities taxonomic and functional variations
in composition after disturbance were followed in a two-dimensional
NMDS ordination space the 30 years monitored. In both cases the NMDS were performed using occurrence-based
(Jaccard) and abundance-based (Bray-Curtis) dissimilarity measures.
Trajectories along time in the plan were reported through the euclidean
distance of successive inventories to the reference inventories in 1989,
i.e. 2-3 years after disturbance, when the uncertainty degree did not exceed
30\% of undetermined trees. To compensate the intrinsinc difference
of initial position among plots, the trajectories corresponded to the euclidian differences along time
with the reference position in 1989. Univariate trajectories of the leaf, stem and
life-history traits were inspected with the community weighted means
(CWM)
\citep{Diaz2007, Garnier2004}. Species seed mass corresponded to
5 classes of increasing mass, seed mass trajectories were therefore
reported as the proportion of each class in the inventories.

@@

\caption{Upper panels, Trajectories of the Simpson taxonomic diversity \textbf{(a)} and Rao functional diversity \textbf{(b)} over 30 years after disturbance. Colors are treaments: green (control), blue (T1), orange (T2), red (T3) with sahded areas the credibility intervals. Lower panels, Relationship between the initial \%AGB removed and Simpson \textbf{(c)} and Rao \textbf{(d)} diversities, 10, 20 and 30 years after disturbance}\label{fig:IDHplot}
\end{figure*}

@@

While both taxonomic and functional composition remained stable in
undisturbed communities (ref to fig), they followed consistent trajectories over time
after disturbance. According to the mapping of functional traits (see appendix I) these
compositional changes corresponded to shifts towards species with more
acquisitive functional strategies, from communities with high average WD
to high average SLA and chlorophyll content. For disturbed communities
the distance of successive inventories to the 1989 reference inventory
followed unimodal trajectories translating cyclic compositional changes
with a recovery of the initial composition (Figure \ref{fig:NMDSplans}).
The maximum dissimilarity with the initial state was positively
correlated to the disturbance treatment for both taxonomic and
functional composition (\(\rho_{spearman}^{taxonomic}=0.91\) and
\(\rho_{spearman}^{functional}=0.96\) respectively) and the time at
maximum was reached around 26 years after disturbance for taxonomic
composition and 22 years for functional composition.

@@

\caption{Plot trajectories in terms of flora composition (left panels \textbf{(a)} and \textbf{(c)}) and functional composition (right panels \textbf{(b)} and \textbf{(d)}) in a the two-dimensional NMDS space. Lower panels (\textbf{(c)} and \textbf{(d)}) represent the euclidean distance to initial condition along the 30 sampled years. Colors are disturbance treatments (green for control, blue for T1,orange for T2 and red for T3) with shaded areas, the credibility intervals}\label{fig:NMDSplans}
\end{figure*}

@@

\caption{Trajectories of the communities weighted means (CWM) over 30 years after disturbance of 4 leaf traits (Leaf thickness, \emph{L\_thickness}, chlorophyll content, \emph{L\_chloro}, toughness, \emph{L\_toughness} and specific area, \emph{SLA}), 2 stem traits (wood specific gravity, \emph{WD}, and bark thickness, \emph{Bark-thick}) and one life history trait (Specific maximum height at adult stage, \emph{Hmax}). Colors are disturbance treatments: green (control), blue (T1), orange (T2), red (T3) with shaded areas the credibility intervals.}\label{fig:CWM}
\end{figure*}

@@

Communities functional redundancy remained stable in control
plots but after disturbance the redundancy trajectories were quite
variable (See appendix I) and apparently independently of the initial
disturbance. Globally after most intense disturbance (plots T2 and T3)
communities redundancy decreased at first place before increasing to
edge, recover or exceed the initial value. *I would suggest to consider only the Functional redundancy restricted to the functional space in the paper*

@@

\caption{Trajectories of functional redundancy within the initial functional space over 30 years after disturbance. Colors are disturbance treatments: green (control), blue (T1), orange (T2), red (T3) with shaded areas the credibility intervals.}(\#fig:RedFun_rest)
\end{figure}

@@

Both communities taxonomic and functional diversities and composition
proved resilient, following similar humped-shaped trajectories starting to return
towards initial values 30 years after disturbance. Communities functional signature is the most
direct link between biodiversity and ecosystem functioning
\citep{Diaz2005} and its resilience meant the recovery of ecosystem
processes in the long term \citep{Guariguata2001}. The resilience of
communities taxonomy, despite the multiple NMDS starting points, suggest the long-term maintenance of the initial
differences in species composition and diversity, in turn suggesting the existence of multiple stable
equilibria as assumed for highly diverse and productive ecosystems
\citep{Chase2003} and the dependency of recovery trajectories on the
initial composition to be recovered
\citep{Hubbell1999, Molino2001, Anderson2007, Baraloto2012a}.

@@

Disturbance intensity poorly predicted the taxonomic richness regardless of the period studied, as
already observed in the Guiana Shield \citep{Baraloto2012a} and in
Bornean tropical forests \citep{Cannon1998}. Contrastingly, the disturbance intensity consistently predicted the
increase in functional diversity after disturbance along the 30 years.
The functional diversity increase paralleled the significant functional shifts towards resource-acquisitive strategies
(sharp increase in the SLA, leaf thickness and bark thickness and
decrease in wood density, leaf toughness and maximum height)
\citep{Westoby1998, Wright2004, Reich2014}. As the old-growth,
pre-disturbance survivor trees proved mirroring the initial communities
\citep{Herault2018} the functional trajectories and taxonomic turnover
rather relied upon the pool of newly recruited trees driven by the
enhanced growth and survival of previously infrequent species and
functional types. Disturbance entailed a reorganization, benefiting to
pioneers and light demanding species, of the typical high dominance
structure of hyperdiverse mature forests. The changes in abiotic
environment and competitive pressure favored pioneers which outcompeted
other species in non limiting light resources but were excluded in
mature forests by long-lived, more resistant and shade tolerant species.

* Not clear from here*
Disturbance trajectories therefore relied upon the environmental niches
made available and filled by species from a restricted functional range
that became dominant and structured the functional characteristics of
the community \citep{Grime1998}. The intermediate disturbance hypothesis
was suitabe to predict communities functional diversity and pool of
dominant species \citep{Molino2001}, but it poorly represented the whole
taxonomic structure that was hampered by the slow recovery of rare
species \citep{Hubbell2001, Chave2004}.

@@

The functional and taxonomic decoupling as well as the rebuttal of the IDH for
communities taxonomic structure were explained by a reorganization of
the functional redundancy within the functional traits space that
implied the incomplete resilience (???) of communities
\citep{Trenbath1999, Elmqvist2003, Diaz2005}.

The functional redundancy restricted to the functional space of the
initial community clearly followed humped shaped trajectories. After
disturbance communities differently occupied the functional space,
following the observed shifts in traits trajectories and functional
composition, and the redundancy in the initial functional space
decreased according to the disturbance intensity. After 30 years the
recovery of the initial redundancy structure was consistent but remained
unachieved. ?????-> It depended on infrequent species specific to old-growth
forests that were functionally and therefore underwent competitive
exclusion limiting functional similarity
\citep{Chave2004, Mayfield2010}. ?????->The persistent alteration of functional
redundancy meant a lower resilience of pre-disturbance communities and
an higher resilience disturbance-specific communities (corresponding to
lower richness and higher dominance of pioneer species) so chances to
observe self-maintained compositional changes in favor of disturbance
resistant species, lianas or epiphytes were increased after disturbance
\citep{Haddad2008, Burslem2000, Martin2013}. ?? Not so long no? -> Communities resilience
depended on the functional redundancy structure that proved long to
recover. This decades-long recovery specifically increased the risks of
loosing cornerstone (????) species, with unexpected ecological consequences
\citep{Jones1994, Chazdon2003a, Diaz2005, Gardner2007}.

@@

Our study defined the decoupled functional and taxonomic trajectories of
tropical rainforests after disturbance demonstrating a rapid functional
recovery but a slower and more variable taxonomic one. Consistently with
the IDH, functional trajectories were driven after disturbance by
enrichment of the system with pioneers and light-demanding species.
????-> The ????following functional shifts entailed a re-organization of
communities functional redundancy that proved long to recover and
explained the more variable and longer taxonomic trajectories as shade
tolerant species of old-growth forests faced functional redundancy and
limiting similarity processes. The resilience of tropical forests then
proved long in the face of quite intense disturbance and did not
preclude the settling of a persistent disturbance-specific community
\citep{Gourlet-Fleury2005}. As the trajectories highlighted the
recruitment processes proved central for communities response to
disturbance and closer focus on demographical drivers of communities
response would clarify the fate of the future forests. ????-> The disturbance
range however stayed within the spectrum of selective logging, with a
forest cover remaining all along the experiment, and the response
mechanisms would probably be much different after harder disturbance.

>>>>

\Abstract{
Understanding the ecological rules underlying the maintenance of
tropical forests biodiversity, structure, functioning and dynamics is
urgent to anticipate their fate in the global changing context. The huge
diversity of tropical forests is often assumed to be shaped by constant
regime of disturbance yielding a diversity peak at intermediate
intensity, but for tropical forests this intermediate disturbance
hypothesis (IDH) remains debated, and this controversy also questions
the extent of communities resilience regarding their functional and
taxonomic facets is also to be clarified. To disentangle the ecological
processes driving communities response to disturbance we analysed the
taxonomic and functional diversity trajectories following a logging and
thinning disturbance gradient. Specifically we examined over 30 years
the trajectories of communities taxonomic richness and evennes and
functional composition, diversity and redundancy based on 7 leaf, stem
and life history traits. Trajectories highlighted communities taxonomic
resilience, maintaining their initial differences in composition, and
functional resilience through fast functional trajectory common to all
communities in the functional space. The IDH consistently predicted
communities functional trajectories according to the disturbance
intensity, but poorly represented their taxonomic trajectories, blurried
by the alteration of the functional redundancy and the recruitment of
infrequent, shade tolerant species slowed by competitive exclusion
processes. Although communities functioning had recovered after 30 years
their taxonomic diversity and composition in infrequent species and
their functional redundancy remained altered. these results acknowledged
the need of decades-long recovery cycles to ensure communities complete
resilience, and questioned communities maintenance after repeated
disturbance.
}

@@

In tropical forest, ecological communities are constantly re-shaped by
natural disturbance events that change both the abiotic environment,
through the fluxes of light, heat and water \citep{Goulamoussene2017},
and the biotic interactions like competition among species
\citep{Chesson2000, Herault2018}. One of the cornerstone of tropical
forest ecology is to understand the processes and drivers of ecosystems
response to disturbance \citep{White2001, Chazdon2003a}. For now, this
has been largely studied through forest structural parameters, rapid and
convenient to measure, as aboveground biomass, tree height or stem
density \citep{Piponiot2016, Rutishauser2016}. These structural
parameters were thereafter sucessfully modeled, giving important
insights into the recovery of ecosystems processes and services
\citep{Herault2018}. However the response of forests diversity is still
unclear, albeit it determines the productivity, stability and
functioning of ecosystems \citep[\citet{Liang2016}]{Tilman2014} and
would be most probably impacted by the changes induced by disturbance
\citep{Baraloto2012a}.

@@

@@

The study site corresponds to a tropical rainforest typical of the
Guiana Shield with a dominance of Fabaceae, Chrysobalanaceae,
Lecythidaceae and Sapotaceae botanical families. In the twelve
experimental plots of the experiment, all trees above 10 cm DBH have
been mapped and measured annually since 1984. Trees are first identified
during inventories with a vernacular name assigned by the forest worker
team, and afterward with a scientific name assigned by botanists during
regular botanical campaigns. In 1984, specific vernacular names are
given to 62 commercial or common species whereas more infrequent ones
were identified under general identifiers only distinguishing trees and
palm trees. From 2003, botanical campaigns have been conducted every 5
to 6 years to identify all trees at the species level but identification
levels still varied among plots and campaigns.

@@

Six functional traits representing leaf economics (leaves thickness,
toughness, total chlorophyll content and specific leaf area, the leaf
area per unit dry mass) and wood economics (wood specific gravity and
bark thickness), and life history traits (maximum specific height and
seed mass) came from the BRIDGE project \footnote{http://www.ecofog.gf/Bridge/}.
Traits values were assessed from a selection of individuals located in
nine permanent plots in French Guiana, including two in Paracou, and
comprised 294 botanical species pertaining to 157 botanical genera.
Missing trait values were filled using multivariate imputation by
chained equation \citep{Mice2011}. Imputations were restricted within
genus, or family when samples were too scarce, in order to account for
the phylogenetic signal of the functional traits. Whenever a species
inventoried was not in the dataset, it was attributed a set of traits
values randomly sampled among species of the same next higher taxonomic
level (same genus or family). As seed mass information corresponds to a
classification into mass classes, no data filling process was applied
and analysis were restricted to the 414 botanical species recorded.

@@

To counter taxonomic uncertainties due to the variability of botanical
identification levels (in space) and protocols (in time), the taxonomic
composition and diversity analysis were conducted at the genus level,
\emph{i.e.} referring to the genus of observed or trialed botanical
names. Trajectories of communities taxonomic and functional variations
in composition after disturbance were followed in a two-dimensional NMDS
ordination space of the 30 years monitored. Two NMDS were conducted to
map either taxonomic flora inventories or communities functional
composition based on the 7 leaf, stem and life history traits (without
seed mass classes). In both cases the NMDS were performed using
occurrence-based (Jaccard) and abundance-based (Bray-Curtis)
dissimilarity measures. Trajectories along time in the plan were
reported through the euclidean distance of successive inventories to the
reference inventories in 1989, \emph{i.e.} 2-3 years after disturbance,
when the uncertainty degree did not exceed 30\% of undetermined trees.
To compensate the intrinsinc difference among plots, the trajectories
corresponded to the euclidean differences along time with the reference
inventory in 1989. Univariate trajectories of the leaf, stem and
life-history traits were also visualized with the community weighted
means (CWM) \citep{Diaz2007, Garnier2004}. Species seed mass
corresponded to 5 classes of increasing mass, seed mass trajectories
were therefore reported as the proportion of each class in the
inventories.

@@

\caption{Upper panels, Trajectories of the Simpson taxonomic diversity \textbf{(a)} and Rao functional diversity \textbf{(b)} over 30 years after disturbance. Colors are treatments: green (control), blue (T1), orange (T2), red (T3) with shaded areas the credibility intervals. Lower panels, Relationship between the initial \%AGB removed and Simpson \textbf{(c)} and Rao \textbf{(d)} diversities 10, 20 and 30 years after disturbance.}\label{fig:IDHplot}
\end{figure*}

@@

While both taxonomic and functional composition remained stable in
undisturbed communities (Figure \ref{fig:NMDSplans}), they followed
consistent trajectories over time after disturbance which revealed
significant compositional changes. According to the mapping of
functional traits (see appendix I) these compositional changes
corresponded to shifts towards species with more acquisitive functional
strategies, from communities with high average WD to high average SLA
and chlorophyll content. For disturbed communities the distance of
successive inventories to the 1989 reference inventory followed unimodal
trajectories translating cyclic compositional changes with a recovery of
the initial composition (Figure \ref{fig:NMDSplans}). The maximum
dissimilarity with the initial state was positively correlated to the
disturbance treatment for both taxonomic and functional composition
(\(\rho_{spearman}^{taxonomic}=0.91\) and
\(\rho_{spearman}^{functional}=0.96\) respectively) and the time at
maximum was reached around 26 years after disturbance for taxonomic
composition and 22 years for functional composition.

@@

\caption{Plot trajectories in terms of flora composition (left panels \textbf{(a)} and \textbf{(c)}) and functional composition (right panels \textbf{(b)} and \textbf{(d)}) in a two-dimensional NMDS space. Lower panels (\textbf{(c)} and \textbf{(d)}) represent the euclidean distance to initial condition along the 30 sampled years. Colors are treatments: green (control), blue (T1), orange (T2), red (T3) with shaded areas the credibility intervals}\label{fig:NMDSplans}
\end{figure*}

@@

\caption{Trajectories of the communities weighted means (CWM) over 30 years after disturbance of 4 leaf traits (Leaf thickness, \emph{L\_thickness}, chlorophyll content, \emph{L\_chloro}, toughness, \emph{L\_toughness} and specific area, \emph{SLA}), 2 stem traits (wood specific gravity, \emph{WD}, and bark thickness, \emph{Bark-thick}) and one life history trait (Specific maximum height at adult stage, \emph{Hmax}). Colors are treatments: green (control), blue (T1), orange (T2), red (T3) with shaded areas the credibility intervals.}\label{fig:CWM}
\end{figure*}

@@

Communities functional remained stable in control plots but after
disturbance the redundancy trajectories were quite variable (See
appendix I) and apparently independently of the initial disturbance.
Globally after most intense disturbance (plots T2 and T3) communities
redundancy decreased at first place before increasing to edge, recover
or exceed the initial value. \textgreater{} restrict to initial
functional space

@@

\caption{TTrajectories of functional redundancy within the initial functional space over 30 years after disturbance. Colors are disturbance treatments: green (control), blue (T1), orange (T2), red (T3) with shaded areas the credibility intervals.}(\#fig:RedFun_rest)
\end{figure}

@@

Both communities taxonomic and functional diversities and composition
proved resilient, following similar humped-shaped trajectories starting
to return towards initial values 30 years after disturbance. Communities
functional signature is the most direct link between biodiversity and
ecosystem functioning \citep{Diaz2005} and its resilience meant the
recovery of ecosystem processes in the long term \citep{Guariguata2001}.
The resilience of communities taxonomy, despite the multiple NMDS
starting points, suggest the long term maintenance of the initial
differences in species composition and diversity. This maintenance in
turn suggest the existence of multiple stable equilibria as assumed for
highly diverse and productive ecosystems \citep{Chase2003} and the
dependency of recovery trajectories on the initial composition to be
recovered \citep{Hubbell1999, Molino2001, Anderson2007, Baraloto2012a}.

@@

Disturbance intensity poorly predicted the taxonomic richness regadless
of the period studied, as already observed in the Guiana Shield
\citep{Baraloto2012a} and in Bornean tropical forests
\citep{Cannon1998}. Contrastingly, the disturbance intensity
consistently predicted the increase in functional diversity after
disturbance along the 30 years. The functional diversity increase
paralleled the significant functional shifts towards
resource-acquisitive strategies (sharp increase in the SLA, leaf
thickness and bark thickness and decrease in wood density, leaf
toughness and maximum height)
\citep{Westoby1998, Wright2004, Reich2014}. As the old-growth,
pre-disturbance survivor trees proved mirroring the initial communities
\citep{Herault2018} the functional trajectories and taxonomic turnover
rather relied upon the pool of newly recruited trees driven by the
enhanced growth and survival of previously infrequent species and
functional types. Disturbance entailed a reorganization, benefiting to
pioneers and light demanding species, of the typical high dominance
structure of hyperdiverse mature forests. The changes in abiotic
environment and competitive pressure favored pioneers which outcompeted
other species in non limiting light resources but were excluded in
mature forests by long-lived, more resistant and shade tolerant species.

\begin{quote}
\begin{quote}
Not clear Disturbance trajectories therefore relied upon the
environmental niches made available and filled by species from a
restricted functional range that became dominant and structured the
functional characteristics of the community \citep{Grime1998}. The
intermediate disturbance hypothesis was suitabe to predict communities
functional diversity and pool of dominant species \citep{Molino2001},
but it poorly represented the whole taxonomic structure that was
hampered by the slow recovery of rare species
\citep{Hubbell2001, Chave2004}.
\end{quote}
\end{quote}

@@

The functional and taxonomic decoupling as well as the rebuttal of the
IDH for communities taxonomic structure were explained by a
reorganization of the functional redundancy within the functional traits
space that implied the \textgreater{}\textgreater{} incomplete
resilience of communities \citep{Trenbath1999, Elmqvist2003, Diaz2005}.

The functional redundancy restricted to the functional space of the
initial community clearly followed humped shaped trajectories. After
disturbance communities differently occupied the functional space,
following the observed shifts in traits trajectories and functional
composition, and the redundancy in the initial functional space
decreased according to the disturbance intensity. After 30 years the
recovery of the initial redundancy structure was consistent but remained
unachieved. \textgreater{}\textgreater{} unclear It depended on
infrequent species specific to old-growth forests that were functionally
and therefore underwent competitive exclusion limiting functional
similarity \citep{Chave2004, Mayfield2010}.

\begin{quote}
\begin{quote}
unclear The persistent alteration of functional redundancy meant a lower
resilience of pre-disturbance communities and an higher resilience
disturbance-specific communities (corresponding to lower richness and
higher dominance of pioneer species) so chances to observe
self-maintained compositional changes in favor of disturbance resistant
species, lianas or epiphytes were increased after disturbance
\citep{Haddad2008, Burslem2000, Martin2013}.
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
not so long Communities resilience depended on the functional redundancy
structure that proved long to recover.
\end{quote}
\end{quote}

\begin{quote}
\begin{quote}
unclear (cornerstone?) This decades-long recovery specifically increased
the risks of loosing cornerstone specie, with unexpected ecological
consequences \citep{Jones1994, Chazdon2003a, Diaz2005, Gardner2007}.
\end{quote}
\end{quote}

@@

Our study defined the decoupled functional and taxonomic trajectories of
tropical rainforests after disturbance demonstrating a rapid functional
recovery but a slower and more variable taxonomic one. Consistently with
the IDH, functional trajectories were driven after disturbance by
enrichment of the system with pioneers and light-demanding species.

\begin{quote}
\begin{quote}
unclear The following functional shifts entailed a re-organization of
communities functional redundancy that proved long to recover and
explained the more variable and longer taxonomic trajectories as shade
tolerant species of old-growth forests faced functional redundancy and
limiting similarity processes.
\end{quote}
\end{quote}

The resilience of tropical forests then proved long in the face of quite
intense disturbance and did not preclude the settling of a persistent
disturbance-specific community \citep{Gourlet-Fleury2005}. As the
trajectories highlighted the recruitment processes proved central for
communities response to disturbance and closer focus on demographical
drivers of communities response would clarify the fate of the future
forests.

\begin{quote}
\begin{quote}
unclear The disturbance range however stayed within the spectrum of
selective logging, with a forest cover remaining all along the
experiment, and the response mechanisms would probably be much different
after harder disturbance.
\end{quote}
\end{quote}
